---
title: "Exploring ImpactGov data"
author: "Scott Brenstuhl"
date: "January 27, 2016"
output:
        md_document:
        variant: markdown_github
---

```{r libraries, alert=FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
# html_document
```

## Citywide Performance

```{r}
#https://data.sfgov.org/api/views/5x94-tptc/rows.csv?accessType=DOWNLOAD
perf.measures.FY14 <- read.csv('../data/Citywide_Performance_Measurement_Annual_Report_FY14_Dataset.csv')

perf.measures.FY13 <- read.csv('../data/Citywide_Performance_Measurement_Annual_Report_FY13_Dataset.csv')
```

How wide of a margin should be given to "On target is positive"
goals?
```{r}
group_by(perf.measures.FY14, Performance.Pattern) %>%
    summarise(count = n())

#subset(perf.measures.FY14, Performance.Pattern == 'On target is positive')
```

Acounting for diffent units might be... interesting depending what we decide to
do.
```{r}
group_by(perf.measures.FY14, Unit.of.Measure) %>%
    summarise(count = n())
```

There might be an opportunity to cross over with 311 data here? Not a ton of
info though.
```{r}
nrow(subset(perf.measures.FY14, Program == '311 CUSTOMER SERVICE CENTER'))
```

```{r}
goal.evaluater <- function(row, actual.col, target.col){
    actual = as.numeric(row[actual.col])
    target = as.numeric(row[target.col])
    if(row['Performance.Pattern'] == 'Below target is positive'){
        actual < target
    }else if(row['Performance.Pattern'] == 'Above target is positive'){
            actual > target
    }else{
        # Checked within 5% is good for now, needs more research
        (target-actual)/target < .05 & (target-actual)/target > -.05
    }
}


perf.measures.FY14$met.goal <- apply(perf.measures.FY14, 1, goal.evaluater, 'FY.2014...Total.Actual.Data', 'FY.2014...Total.Target.Data')


perf.measures.FY13$met.goal <- apply(perf.measures.FY13, 1, goal.evaluater, 'FY.2013...Total.Actual.Data', 'FY.2013...Total.Target.Data')

success.by.pattern <-table(perf.measures.FY14[c('met.goal', 'Performance.Pattern')])
success.by.pattern
apply(success.by.pattern, 2, function(col){col[2]/(col[1]+col[2])})

```

```{r echo=FALSE}
bar.order <- names(sort(table(perf.measures.FY14$Dept),decreasing = TRUE))

ggplot(perf.measures.FY14, aes(x = Dept, fill = met.goal)) +
    geom_bar() +
    scale_x_discrete(limits = bar.order) +
    labs(title= "Performance Measure by Department FY14") +
    theme(legend.position="top")+
    theme(axis.text.x = element_text(angle = -45, hjust = 0))

ggplot(perf.measures.FY13, aes(x = Dept, fill = met.goal)) +
    geom_bar() +
    scale_x_discrete(limits = bar.order) +
    labs(title= "Performance Measure by Department FY13") +
    theme(legend.position="none")+
    theme(axis.text.x = element_text(angle = -45, hjust = 0))

unique(subset(perf.measures.FY14, select = c(Dept,Department)))
```

Can't forget there are programs that didnt have measures.


##Budget

```{r}
budget <- read.csv('../data/Budget.csv')
```

Most of the departments are also present in peformance measures FY14
```{r}
budget.depts <- unique(budget$Department.Code)
perf.depts <- unique(perf.measures.FY14$Dept)
length(intersect(budget.depts, perf.depts))

# the only value in perf.depts not in budget are blanks.
# So every department has budget entries!
setdiff(perf.depts, intersect(budget.depts, perf.depts))

no.perf <- setdiff(budget.depts,perf.depts)

no.perf.budget <- subset(budget, Department.Code %in% no.perf)

# Do not have matching FY14 performance values
unique(no.perf.budget$Department)
```

will soon be relasing all KPI for the city
Look at baltimore daily report card
sf government barameter through city controlers office. 
data for good exchange

Look at city controler's office website. City performance section

Optimal distribution of ambulances 

Rail and Bus

Affordable public housing (how do HOA fees affect people)

```{r}

dept.budget.2014 <- subset(budget,Fiscal.Year == 2014) %>%
    group_by(Department.Code) %>%
    summarise(sum(Amount))

dept.budget.2013 <- subset(budget,Fiscal.Year == 2013) %>%
    group_by(Department.Code) %>%
    summarise(sum(Amount))

dept.budget.2013
merge(dept.budget.2013, dept.budget.2014, by = 'Department.Code')
?merge
head(budget)

sort(unique(budget$Program))
         
rail.bus <- subset(budget, Program =='Rail & Bus Services') %>%
    group_by(Character, Fiscal.Year) %>%
    summarise('amount'= sum(Amount)) %>%
    arrange(desc(amount))


    

names(budget)
head(budget[budget$Program == 'Rail & Bus Services',])

```