---
title: "Exploring ImpactGov data"
author: "Scott Brenstuhl"
date: "January 27, 2016"
output:
        md_document:
        variant: markdown_github
---

```{r libraries, alert=FALSE, message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
# html_document
```

## Citywide Performance

```{r}
#https://data.sfgov.org/api/views/5x94-tptc/rows.csv?accessType=DOWNLOAD
perf.measures.FY14 <- read_csv('../data/Citywide_Performance_Measurement_Annual_Report_FY14_Dataset.csv')

perf.measures.FY13 <- read_csv('../data/Citywide_Performance_Measurement_Annual_Report_FY13_Dataset.csv')
```

How wide of a margin should be given to "On target is positive"
goals?
```{r}
group_by(perf.measures.FY14, Performance.Pattern) %>%
    summarise(count = n())

#subset(perf.measures.FY14, Performance.Pattern == 'On target is positive')
```

Acounting for diffent units might be... interesting depending what we decide to
do.
```{r}
group_by(perf.measures.FY14, Unit.of.Measure) %>%
    summarise(count = n())
```

There might be an opportunity to cross over with 311 data here? Not a ton of
info though.
```{r}
nrow(subset(perf.measures.FY14, Program == '311 CUSTOMER SERVICE CENTER'))
```

```{r}
goal.evaluater <- function(row, actual.col, target.col){
    actual = as.numeric(row[actual.col])
    target = as.numeric(row[target.col])
    if(row['Performance.Pattern'] == 'Below target is positive'){
        actual < target
    }else if(row['Performance.Pattern'] == 'Above target is positive'){
            actual > target
    }else{
        # Checked within 5% is good for now, needs more research
        (target-actual)/target < .05 & (target-actual)/target > -.05
    }
}


perf.measures.FY14$met.goal <- apply(perf.measures.FY14, 1, goal.evaluater, 'FY.2014...Total.Actual.Data', 'FY.2014...Total.Target.Data')


perf.measures.FY13$met.goal <- apply(perf.measures.FY13, 1, goal.evaluater, 'FY.2013...Total.Actual.Data', 'FY.2013...Total.Target.Data')

success.by.pattern <-table(perf.measures.FY14[c('met.goal', 'Performance.Pattern')])
success.by.pattern
apply(success.by.pattern, 2, function(col){col[2]/(col[1]+col[2])})

```

```{r echo=FALSE}
bar.order <- names(sort(table(perf.measures.FY14$Dept),decreasing = TRUE))

ggplot(perf.measures.FY14, aes(x = Dept, fill = met.goal)) +
    geom_bar() +
    scale_x_discrete(limits = bar.order) +
    labs(title= "Performance Measure by Department FY14") +
    theme(legend.position="top")+
    theme(axis.text.x = element_text(angle = -45, hjust = 0))

ggplot(perf.measures.FY13, aes(x = Dept, fill = met.goal)) +
    geom_bar() +
    scale_x_discrete(limits = bar.order) +
    labs(title= "Performance Measure by Department FY13") +
    theme(legend.position="none")+
    theme(axis.text.x = element_text(angle = -45, hjust = 0))

unique(subset(perf.measures.FY14, select = c(Dept,Department)))
```

Can't forget there are programs that didnt have measures.


##Budget

```{r}
budget <- read.csv('../data/Budget.csv')
```

Most of the departments are also present in peformance measures FY14
```{r}
budget.depts <- unique(budget$Department.Code)
perf.depts <- unique(perf.measures.FY14$Dept)
length(intersect(budget.depts, perf.depts))

# the only value in perf.depts not in budget are blanks.
# So every department has budget entries!
setdiff(perf.depts, intersect(budget.depts, perf.depts))

no.perf <- setdiff(budget.depts,perf.depts)

no.perf.budget <- subset(budget, Department.Code %in% no.perf)

# Do not have matching FY14 performance values
unique(no.perf.budget$Department)
```


```{r}

dept.budget.2014 <- subset(budget,Fiscal.Year == 2014) %>%
    group_by(Department.Code) %>%
    summarise(sum(Amount))

dept.budget.2013 <- subset(budget,Fiscal.Year == 2013) %>%
    group_by(Department.Code) %>%
    summarise(sum(Amount))

change.13.to.14 <- merge(dept.budget.2013,
                         dept.budget.2014, by = 'Department.Code')

 

```

Figure out how to rule out iner-departmental spending

```{r}
neg.rev.14 <- subset(budget, Amount < 0 & Fiscal.Year == 2014 & Revenue.or.Spending == 'Revenue')

neg.spend.14 <- subset(budget, Amount < 0 & Fiscal.Year == 2014 & Revenue.or.Spending == 'Spending')

sum(neg.rev.14$Amount)
sum(neg.spend.14$Amount)

#figure out if Object is a better breakdown

var <- subset(budget.14, Object == 'Recovery for Svcs to AAO Funds')

serv <- subset(budget.14, Character == 'Services of Other Depts')

uses.14 <- subset(budget,  Amount < 0 & 
                         Fiscal.Year == 2014 & 
                         Character == 'Transfer Adjustments-Uses')

#no match, amount == 9864899


sources.14 <- subset(budget, Amount < 0 & 
                         Fiscal.Year == 2014 & 
                         Character ==  'Transfer Adjustments-Sources')
uses.13 <- subset(budget,  Amount < 0 & 
                         Fiscal.Year == 2013 & 
                         Character == 'Transfer Adjustments-Uses')

sources.13 <- subset(budget, Amount < 0 & 
                         Fiscal.Year == 2013 & 
                         Character ==  'Transfer Adjustments-Sources')

# looks close but not exact
sum(uses.14$Amount)
sum(sources.14$Amount)
sum(uses.13$Amount)
sum(sources.13$Amount)

# uses has two less
table(uses.14$Amount < 0)
table(sources.14$Amount < 0)

table(table(uses.14$Department) == table(sources.14$Department))
table(table(uses.13$Department) == table(sources.13$Department))
```

It looks likely that Airport Commission and Port have a misslabled expense in 14.
13's numbers are similar Airport Commision, Port, and Fire Department are off.

Figuring out the breakdown 

```{r echo = FALSE}
spend.rev <- read.csv('../data/Spending_And_Revenue.csv')

budget.14 <- subset(budget, Fiscal.Year == 2014)

spend.rev.14 <- subset(spend.rev, Fiscal.Year == 2014)

unique(spend.rev.14) == unique(budget.14)

budget.14 

```

```{r echo = FALSE}
taxes <- budget[grep('Tax',budget$Sub.object),]
taxes <- subset(taxes, Revenue.or.Spending == 'Revenue')
var2<- group_by(taxes, Fiscal.Year, Organization.Group) %>%
    summarise(total = sum(Amount)/1000000)

ggplot(var2, aes(x=Fiscal.Year, y=total))+
    facet_grid(.~Organization.Group) +
    geom_point()

```

```{r echo=FALSE}
View(budget.14)

```

Are over spenders more likely to meet their own goals
